\chapter{Simple skew braces}
\label{Sbraces}


As we have seen in Chapter \ref{Isolutions}, the classification of finite involutive solutions to the YBE is reduced to the classification of the finite skew braces of abelian type. A first step in the classification of the finite skew braces of abelian type is the classification of the finite simple skew braces of abelian type.

In this chapter we study some constructions of finite simple skew braces of abelian type. 

\section{Finite simple skew braces}

\begin{definition}
\index{Skew brace!simple}
A non-zero skew brace $B$ is \emph{simple} if $\{0\}$ and $B$ are the only ideals of $B$.
\end{definition}


\begin{example}
\index{Skew brace!simple trivial}
Let $G$ be a simple additive group. Then the trivial skew brace $(G,+,+)$, is a simple skew brace. These are the simple trivial skew braces.
\end{example}

Thus the classification of the finite simple trivial skew braces is equivalent to the classification of the finite simple groups.
In particular, the finite simple trivial skew braces of abelian type are the trivial skew braces of prime cardinality. 

By Theorem \ref{thm:add_nilpotent}, the multiplicative group of a finite skew brace of nilpotent type is solvable. Let $B$ be a non-zero finite skew brace of nilpotent type. Let $p_1,\dots ,p_k$ be the distinct prime divisors of $|B|$. Let $B_{p_i}$ be the Sylow $p_i$-subgroup of the additive group of $B$, for $i=1,\dots ,k$. Note that $B_{p_i}$ is a left ideal of $B$, and the multiplicative groups of these left ideals $B_{p_i}$ form a Sylow system of the multiplicative group of $B$. Thus
$$B=\prod_{i=1}^kB_{p_i}.$$
Note that for $a\in B_{p_i}$, $b\in B_{p_j}$, where $i\neq j$, there exist unique $c\in B_{p_j}$ and $d\in B_{p_i}$ such that $a\circ b=c\circ d$. On the other hand, 
$$a\circ b=\lambda_a(b)\circ \lambda_a(b)'\circ a\circ b=\lambda_a(b)\circ (b'+\lambda_{b'}(a')-b')'.$$
Hence $c=\lambda_a(b)$ and $d=(b'+\lambda_{b'}(a')-b')'$.
    


\begin{lemma}
    \label{lem:nilpotentsimple}
    Let $B$ be a finite simple skew brace of nilpotent type with nilpotent multiplicative group. Then $B$ is a trivial skew brace of prime order. 
\end{lemma}

\begin{proof}
    Let $p$ be a prime divisor of $|B|$. Let $B_p$ be the Sylow $p$-subgroup of the additive group of $B$. We know that $B_p$ is a left ideal of $B$. In particular, $B_p$ also is a Sylow $p$-subgroup of the multiplicative group of $B$. Since the multiplicative group of $B$ is nilpotent, we have that $B_p$ is an ideal of $B$. Since $B$ is simple, $B=B_p$. Let $G$ be the multiplicative group of $B$ and let $A$ be the additive group of $B$. Then the semidirect product $H=A\rtimes G$ via the lambda map is a $p$-group, and thus it is nilpotent. Let $a\in A$ and $g\in G$. We have that 
    $$(a,0)(0,g)(-a,0)(0,g')=(a,g)(-a,g')=(a-\lambda_g(a),0)=(-(g*a),0).$$
    Therefore $[\{0\}\times G,A\times\{0\}]=B*B\times \{0\}$. Since $H$ is nilpotent, it is clear that $B*B\neq B$. Since $B*B$ is an ideal of $B$ and $B$ is simple, it follows that $B*B=\{0\}$. Hence $B$ is a trivial skew brace. Since $B$ is simple, $|B|=p$.   
\end{proof}

The above result shows that if $B$ is a finite non-trivial simple skew brace of abelian type, 
then there exist two distinct prime divisors of $|B|$. To understand the structure of $B$ we need to study the structure of the left ideals $B_p$, where $B_p$ is the Sypow $p$-subgroup of $(B,+)$, and the restriction of the lambda map to $B_p$ acting on $B_q$, where $p,q$ are distinct prime divisors of $|B|$.


Let $B$ be a finite simple skew brace of abelian type of order $p^nq^m$ for distinct primes $p,q$ and positive integers $n,m$. Let $B_p$ and $B_q$ be the Sylow $p$-subgroup and the Sylow $q$-subgroup of the additive group of $B$ respectively. Let $\alpha\colon B_q\rightarrow \Aut(B_p,+)$ and $\beta\colon B_p\rightarrow \Aut(B_q,+)$ be restictions of the lambda map of $B$. Let $a\in B_p$ and $b\in B_q$ and suppose that $(a,b)\neq (0,0)$. Since $B$ is simple, we have that the ideal generated by $a+b$ is $B$. In particular,  if $a\neq 0$, then the ideal generated by $a$ is $B$. Let $I_0$ be a minimal nonzero ideal of $B_p$. Clearly $I_0$ is not an ideal of $B$. This means that there exist $d\in B_q$ and $c\in I_0$ such that $c*d=(\lambda_c-\id)(d)=(\beta_c-id)(d)\neq 0$. Let 
$J_0$ be the ideal of $B_q$ generated by $\{ (\beta_c-id)(d) : c\in I_0, \; d\in B_q\}$. Let $I_1$ be the ideal of $B_p$ generated by 
$$I_0\cup \{(\alpha_d-id)(c) : d\in J_0,\; c\in B_p\}.$$ 
Let $J_1$ be the ideal of $B_q$ generated by 
$$J_0\cup \{(\beta_c-id)(d) : c\in I_1,\; d\in B_q\}.$$
For $k>1$ let  $I_k$ be the ideal of $B_p$ generated by 
$$I_{k-1}\cup \{(\alpha_d-id)(c) : d\in J_{k-1},\; c\in B_p\},$$ and let $J_K$ be the ideal of $B_q$ generated by 
$$J_{k-1}\cup \{(\beta_c-id)(d) : c\in I_k,\; d\in B_q\}.$$
Then there exists a positive integer $k$ such that $I_k=B_p$ and $J_k=B_q$. 

We shall study the main constructions of the finite simple braces 
of abelian type of order $p^nq^m$ for distinct primes $p,q$ and positive integers $n,m$.
  
\section{Matched product of skew braces} 

Let $B$ be a skew brace. Suppose that $(B,+)$ is the direct product of two left ideals $B_1$ and $B_2$ of $B$.
Note that for $a,x\in B_1$ and $b,y\in B_2$, we have 
\begin{eqnarray*}
	\lambda_{a+b}(x+y)&=&\lambda_{a+b}(x)+\lambda_{a+b}(y)\\
	&=&\lambda_{a}\lambda_{\lambda^{-1}_{a}(b)}(x)+\lambda_{a}\lambda_{\lambda^{-1}_{a}(b)}(y)\\
	&=&\lambda_{b}\lambda_{\lambda^{-1}_{b}(a)}(x)+\lambda_{b}\lambda_{\lambda^{-1}_{b}(a)}(y),
\end{eqnarray*}	
where the last equality follows because $a+b=b+a$.

Let $\alpha\colon (B_2,\circ)\rightarrow\Aut(B_1,+)$ and $\beta\colon (B_1,\circ)\rightarrow\Aut(B_2,+)$ be the restrictions of the lambda map, i.e. 
$$\alpha_b(x)=\lambda_b(x)\mbox{ and }\beta_a(y)=\lambda_a(y),$$
where $\alpha_b=\alpha(b)$ and $\beta_a=\beta(a)$, for all $a,x\in B_1$ and $b,y\in B_2$. Then
$$\alpha_b\lambda^{(1)}_{\alpha^{-1}_b(a)}=\lambda^{(1)}_a\alpha_{\beta^{-1}_a(b)}$$
and	
$$\beta_a\lambda^{(2)}_{\beta^{-1}_a(b)}=\lambda^{(2)}_b\beta_{\alpha^{-1}_b(a)},$$
for all $a\in B_1$ and $b\in B_2$, where $\lambda^{(1)}$ is the lambda map of $B_1$ and $\lambda^{(2)}$ is the lambda map of $B_2$.

\begin{definition}\label{def:matchedpair}
	\index{matched pair}
	Let $B_1$ and $B_2$ be two skew braces. Let $\alpha\colon (B_2,\circ)\rightarrow \Aut(B_1,+)$ and $\beta\colon (B_1,\circ)\rightarrow \Aut(B_2,+)$ be group homomorphisms. We say that $(B_1,B_2,\alpha,\beta)$ is a \emph{matched pair} of skew braces if
	$$\alpha_b\lambda^{(1)}_{\alpha^{-1}_b(a)}=\lambda^{(1)}_a\alpha_{\beta^{-1}_a(b)}$$
	and	
	$$\beta_a\lambda^{(2)}_{\beta^{-1}_a(b)}=\lambda^{(2)}_b\beta_{\alpha^{-1}_b(a)},$$
	where $\alpha_b=\alpha(b)$ and $\beta_a=\beta(a)$, for all $a\in B_1$ and $b\in B_2$, and where $\lambda^{(1)}$ is the lambda map of $B_1$ and $\lambda^{(2)}$ is the lambda map of $B_2$.	
\end{definition}

\begin{lemma} \label{lemma:lambda}
	Let $(B,+)$ be a group (not necessarily abelian). Let
	$$\lambda\colon B\rightarrow \Aut(B.+),\quad a\mapsto \lambda_a,$$ 
	be a map such that
	$\lambda_a\lambda_b=\lambda_{a+\lambda_a(b)}$, for all $a,b\in B$. Then $B$ with the sum $+$ and the product defined by $a\circ b:=a+\lambda_a(b)$, for all $a,b\in B$, is a skew brace.	
\end{lemma}	

\begin{proof}
	Note that, if $0$ is the neutral element of $(B,+)$, then $\lambda_0\lambda_0=\lambda_{\lambda_0(0)}=\lambda_0$. Hence
	$\lambda_0=\id$. Since $a\circ 0=a+\lambda_a(0)=a+0=a$ and $0\circ a=0+\lambda_0(a)=a$, we have that $0$ is the neutral element of $(B,\circ)$. Since $\lambda_a\lambda_{\lambda^{-1}_{a}(-a)}=\lambda_{a-a}=\lambda_0=\id$, we get that $\lambda^{-1}_a=\lambda_{\lambda^{-1}_a(-a)}$. Now $a\circ \lambda^{-1}_{a}(-a)=a-a=0$ and $\lambda^{-1}_a(-a)\circ a=\lambda^{-1}_a(-a)+\lambda_{\lambda^{-1}_a(-a)}(a)=-\lambda^{-1}_a(a)+\lambda^{-1}_{a}(a)=0$. Hence $\lambda^{-1}_a(-a)$ is the symmetric element of $a$ in $(B,\circ)$. We shall check the associativity property:
	\begin{eqnarray*}
		a\circ (b\circ c)&=&a+\lambda_a(b+\lambda_b(c))=a+\lambda_a(b)+\lambda_a(\lambda_b(c))\\
		&=&a+\lambda_a(b)+\lambda_{a+\lambda_a(b)}(c)=(a\circ b)+\lambda_{a\circ b}(c)\\
		&=&(a\circ b)\circ c.
	\end{eqnarray*}
Hence $(B,\circ)$ is a group. Note that
\begin{eqnarray*}
	a\circ (b+c)&=&a+\lambda_a(b+c)=a+\lambda_a(b)+\lambda_a(c)\\
	&=&a+\lambda_a(b)-a+a+\lambda_a(c)=a\circ b-a+a\circ c.
\end{eqnarray*}
Therefore $(B,+,\circ)$ is a skew brace, and the result follows.
\end{proof}

\begin{theorem}\label{thm:matchedproduct}
	Let $(B_1,B_2,\alpha,\beta)$ be a matched pair of skew braces. Then $B_1\times B_2$ is a skew brace with sum given by
	$$(a,b)+(c,d)=(a+c.b+d)$$
and lambda map given by
$$\lambda_{(a,b)}(c,d):=\left( \lambda^{(1)}_{a}\alpha_{\beta^{-1}_a(b)}(c), \lambda^{(2)}_{b}\beta_{\alpha^{-1}_b(a)}(d)\right).$$
Moreover $B_1\times \{0\}$ and $\{0\}\times B_2$ are left ideals isomorphic to $B_1$ and $B_2$ as skew braces respectively.
This skew brace is \index{matched product} the \emph{matched product} of $B_1$ and $B_2$, and it is denoted by $B_1\bowtie B_2$.

Conversely, if $B$ is a skew brace such that $(B,+)$ is the direct product of two left ideals $L_1$ and $L_2$ of $B$, then $(L_1,L_2,\alpha,\beta)$ is a matched pair of left braces, where $\alpha$ and $\beta$ are the restrictions of the lambda map of $B$, and $B$  is isomorphic to the corresponding matched product $L_1\bowtie L_2$.   	
\end{theorem}	

\begin{proof}
	To prove the first part, by Lemma \ref{lemma:lambda}, it is enough to check that
	$$\lambda_{(a,b)}\lambda_{(c,d)}=\lambda_{(a,b)+\lambda_{(a,b)}(c,d)},$$
	for all $a,c\in B_1$ and $b,d\in B_2$. Let $x\in B_1$ and $y\in B_2$. We have that
	\begin{eqnarray*}
		\lambda_{(a,b)}\lambda_{(c,d)}(x,y)&=&\lambda_{(a,b)}\left( \lambda^{(1)}_{c}\alpha_{\beta^{-1}_{c}(d)}(x), \lambda^{(2)}_{d}\beta_{\alpha^{-1}_{d}(c)}(y)\right)\\
		&=&\left( \lambda^{(1)}_{a}\alpha_{\beta^{-1}_{a}(b)}\lambda^{(1)}_{c}\alpha_{\beta^{-1}_{c}(d)}(x), \lambda^{(2)}_{b}\beta_{\alpha^{-1}_{b}(a)}\lambda^{(2)}_{d}\beta_{\alpha^{-1}_{d}(c)}(y)\right)
	\end{eqnarray*}	
and
	\begin{eqnarray*}
	\lambda_{(a,b)+\lambda_{(a,b)}(c,d)}(x,y)&=&\lambda_{(a+\lambda^{(1)}_{a}\alpha_{\beta^{-1}_{a}(b)}(c),b+\lambda^{(2)}_{b}\beta_{\alpha^{-1}_{b}(a)}(d))}(x,y)\\
	&=&\left( \lambda^{(1)}_{a+\lambda^{(1)}_{a}\alpha_{\beta^{-1}_{a}(b)}(c)}\alpha_{\beta^{-1}_{a+\lambda^{(1)}_{a}\alpha_{\beta^{-1}_{a}(b)}(c)}(b+\lambda^{(2)}_{b}\beta_{\alpha^{-1}_{b}(a)}(d))}(x),\right.\\
	&&\quad\left. \lambda^{(2)}_{b+\lambda^{(2)}_{b}\beta_{\alpha^{-1}_{b}(a)}(d)}\beta_{\alpha^{-1}_{b+\lambda^{(2)}_{b}\beta_{\alpha^{-1}_{b}(a)}(d)}(a+\lambda^{(1)}_{a}\alpha_{\beta^{-1}_{a}(b)}(c))}(y)\right).
\end{eqnarray*}	
Thus, it is enough to check the following equalities:
\begin{equation}\label{eq:matched1}
	\lambda^{(1)}_{a}\alpha_{\beta^{-1}_{a}(b)}\lambda^{(1)}_{c}\alpha_{\beta^{-1}_{c}(d)}=\lambda^{(1)}_{a+\lambda^{(1)}_{a}\alpha_{\beta^{-1}_{a}(b)}(c)}\alpha_{\beta^{-1}_{a+\lambda^{(1)}_{a}\alpha_{\beta^{-1}_{a}(b)}(c)}(b+\lambda^{(2)}_{b}\beta_{\alpha^{-1}_{b}(a)}(d))}
\end{equation}
and
\begin{equation}\label{eq:matched2}
	\lambda^{(2)}_{b}\alpha_{\alpha^{-1}_{b}(a)}\lambda^{(2)}_{d}\beta_{\alpha^{-1}_{d}(c)}=\lambda^{(2)}_{b+\lambda^{(2)}_{b}\beta_{\alpha^{-1}_{b}(a)}(d)}\beta_{\alpha^{-1}_{b+\lambda^{(2)}_{b}\beta_{\alpha^{-1}_{b}(a)}(d)}(a+\lambda^{(1)}_{a}\alpha_{\beta^{-1}_{a}(b)}(c))}.
\end{equation}
We shall prove (\ref{eq:matched1}). The equality (\ref{eq:matched2}) is proved similarly. 
We have
\begin{eqnarray*}
	\lefteqn{\lambda^{(1)}_{a+\lambda^{(1)}_{a}\alpha_{\beta^{-1}_{a}(b)}(c)}\alpha_{\beta^{-1}_{a+\lambda^{(1)}_{a}\alpha_{\beta^{-1}_{a}(b)}(c)}(b+\lambda^{(2)}_{b}\beta_{\alpha^{-1}_{b}(a)}(d))}}\\
	&=& \lambda^{(1)}_{a\circ \alpha_{\beta^{-1}_{a}(b)}(c)}\alpha_{\beta^{-1}_{a\circ \alpha_{\beta^{-1}_{a}(b)}(c)}(b+\lambda^{(2)}_{b}\beta_{\alpha^{-1}_{b}(a)}(d))}\\
	&=& \lambda^{(1)}_{a\circ \alpha_{\beta^{-1}_{a}(b)}(c)}\alpha_{\beta^{-1}_{a\circ \alpha_{\beta^{-1}_{a}(b)}(c)}(b)+\beta^{-1}_{a\circ \alpha_{\beta^{-1}_{a}(b)}(c)}(\lambda^{(2)}_{b}\beta_{\alpha^{-1}_{b}(a)}(d))}\\
	&=& \lambda^{(1)}_{a\circ \alpha_{\beta^{-1}_{a}(b)}(c)}\alpha_{\beta^{-1}_{a\circ \alpha_{\beta^{-1}_{a}(b)}(c)}(b)+\lambda^{(2)}_{\beta^{-1}_{a\circ \alpha_{\beta^{-1}_{a}(b)}(c)}(b)}\beta^{-1}_{\alpha^{-1}_b(a\circ \alpha_{\beta^{-1}_a(b)}(c))}\beta_{\alpha^{-1}_{b}(a)}(d)}\\
	&=& \lambda^{(1)}_{a\circ \alpha_{\beta^{-1}_{a}(b)}(c)}\alpha_{\beta^{-1}_{a\circ \alpha_{\beta^{-1}_{a}(b)}(c)}(b)\circ \beta^{-1}_{\alpha^{-1}_b(a\circ \alpha_{\beta^{-1}_a(b)}(c))}\beta_{\alpha^{-1}_{b}(a)}(d)}\\
	&=& \lambda^{(1)}_{a\circ \alpha_{\beta^{-1}_{a}(b)}(c)}\alpha_{\beta^{-1}_{a\circ \alpha_{\beta^{-1}_{a}(b)}(c)}(b)\circ \beta^{-1}_{\alpha^{-1}_b(a+\lambda^{(1)}_a\alpha_{\beta^{-1}_a(b)}(c))}\beta_{\alpha^{-1}_{b}(a)}(d)}\\
	&=& \lambda^{(1)}_{a\circ \alpha_{\beta^{-1}_{a}(b)}(c)}\alpha_{\beta^{-1}_{a\circ \alpha_{\beta^{-1}_{a}(b)}(c)}(b)\circ \beta^{-1}_{\alpha^{-1}_b(a+\alpha_b\lambda^{(1)}_{\alpha^{-1}_b(a)}(c))}\beta_{\alpha^{-1}_{b}(a)}(d)}\\
	&=& \lambda^{(1)}_{a\circ \alpha_{\beta^{-1}_{a}(b)}(c)}\alpha_{\beta^{-1}_{a\circ \alpha_{\beta^{-1}_{a}(b)}(c)}(b)\circ \beta^{-1}_{\alpha^{-1}_b(a)+\lambda^{(1)}_{\alpha^{-1}_b(a)}(c)}\beta_{\alpha^{-1}_{b}(a)}(d)}\\
	&=& \lambda^{(1)}_{a\circ \alpha_{\beta^{-1}_{a}(b)}(c)}\alpha_{\beta^{-1}_{a\circ \alpha_{\beta^{-1}_{a}(b)}(c)}(b)\circ \beta^{-1}_{\alpha^{-1}_b(a)\circ c}\beta_{\alpha^{-1}_{b}(a)}(d)}\\
	&=& \lambda^{(1)}_{a\circ \alpha_{\beta^{-1}_{a}(b)}(c)}\alpha_{\beta^{-1}_{a\circ \alpha_{\beta^{-1}_{a}(b)}(c)}(b)\circ \beta^{-1}_{c}(d)}\\
	&=& \lambda^{(1)}_{a}\lambda^{(1)}_{\alpha_{\beta^{-1}_{a}(b)}(c)}\alpha_{\beta^{-1}_{\alpha_{\beta^{-1}_{a}(b)}(c)}\beta^{-1}_a(b)}\alpha_{\beta^{-1}_{c}(d)}\\
	&=& \lambda^{(1)}_{a}\alpha_{\beta^{-1}_a(b)}\lambda^{(1)}_{c}\alpha_{\beta^{-1}_{c}(d)}.
\end{eqnarray*}
This finishes the proof of the first part.

The second part follows easily by the argument above Definition \ref{def:matchedpair}.
\end{proof}
The matched product of two skew braces can be generalized to the matched product of more than two skew braces following the case of skew braces of abelian type (see \cite{MR3812099} for the matched product of several skew braces of abelian type). 

Let $B$ be a finite skew brace of nilpotent type. Let $B_1,\dots ,B_n$ be the Sylow subgroups of $(B,+)$. Then $(B.+)$ is the direct product of these Sylow subgroups and, moreover, $B_1,\dots ,B_n$ are left ideals of $B$. Thus $B$ is the matched product of $B_1,\dots ,B_n$. In particular, every simple non-trivial skew brace of abelian type is the matched product of two or more skew braces of abelian type of order a power of a prime. But, in fact all the known simple non-trivial skew braces of abelian type are constructed using another kind of product, the so called asymmetric product.

\section{Asymmetric product of skew braces of abelian type}

Let $S$ and $T$ be two (additive) abelian groups. Recall that a (normalized)
\emph{symmetric $2$-cocyle} \index{symmetric $2$-cocyle} on $T$ with values in $S$ is a map $b\colon
T\times T\to  S$ such that
\begin{enumerate}
	\item[(i)] $b(0,0)=0$;
	\item[(ii)] $b(t_1,t_2)=b(t_2,t_1)$;
	\item[(iii)] $b(t_1+t_2,t_3)+b(t_1,t_2)=b(t_1,t_2+t_3)+b(t_2,t_3)$, for all $t_1,t_2,t_3\in T$.
\end{enumerate}
As a consequence, we get that $b(t,0)=b(0,t)=0$, for all $t\in T$.

\begin{theorem}\label{ccs}
	Let $T$ and $S$ be two skew braces of abelian type. Let $b \colon T\times T
	\to S$ be a symmetric $2$-cocycle on $(T, +)$ with
	values in $(S, +)$, and let $\alpha \colon (S,\circ)\to \Aut(T,+,\circ)$ be a homomorphism of groups such that
	\begin{eqnarray}\label{cond3} && s\circ b(t_2, t_3) + b(t_1\circ \alpha_s(t_2 + t_3),t_1)=
		b(t_1\circ\alpha_s(t_2), t_1\circ\alpha_s(t_3))+ s,\end{eqnarray}
	where $\alpha_s=\alpha(s)$, for all $s\in S$ and $t_1, t_2, t_3\in T$. Then the addition and
	multiplication over $T\times S$ given by
	$$
	(t_1,s_1)+(t_2,s_2)=(t_1+t_2,~s_1+s_2+b(t_1,t_2)),
	$$
	$$
	(t_1,s_1)\circ (t_2,s_2)=(t_1\circ \alpha_{s_1}(t_2),~s_1\circ s_2),
	$$
	define a structure of skew brace of abelian type on $T\times S$. We call this skew
	brace the \emph{asymmetric product} \index{asymmetric product} of $T$ by $S$ (via $b$ and $\alpha$)
	and denote it by  $T\rtimes_\circ S$.
\end{theorem}

\begin{proof}
	It is easy to check that $(T\times S,+)$ is an abelian group and that $(T\times S,\circ)$ is a group. In fact, 
	$(T\times S,+)$ is the extension of the abelian groups $S$ and $T$ via the $2$-cocycle $b$, and $(T\times S, \circ)$ is the semidirect product $T\rtimes_{\alpha}S$ of $T$ by $S$ via $\alpha$. Thus it is enough to check that
	$$(t_1,s_1)\circ ((t_2,s_2)+(t_3,s_3))=(t_1,s_1)\circ (t_2,s_2)-(t_1,s_1)+(t_1,s_1)\circ (t_3,s_3).$$
	We have that
	\begin{eqnarray*}
		\lefteqn{(t_1,s_1)\circ (t_2,s_2)-(t_1,s_1)+(t_1,s_1)\circ (t_3,s_3)}\\
		&=&(t_1\circ\alpha_{s_1}(t_2),s_1\circ s_2)-(t_1,s_1)+(t_1\circ\alpha_{s_1}(t_3),s_1\circ s_3)\\
		&=&(t_1\circ\alpha_{s_1}(t_2)-t_1,s_1\circ s_2-s_1-b(t_1\circ\alpha_{s_1}(t_2)-t_1,t_1))+(t_1\circ\alpha_{s_1}(t_3),s_1\circ s_3)\\
		&=&(t_1\circ\alpha_{s_1}(t_2)-t_1+t_1\circ\alpha_{s_1}(t_3),s_1\circ s_2-s_1-b(t_1\circ\alpha_{s_1}(t_2)-t_1,t_1)+s_1\circ s_3\\
		&&\qquad+b(t_1\circ\alpha_{s_1}(t_2)-t_1,t_1\circ\alpha_{s_1}(t_3)))\\
        &=&(t_1\circ(\alpha_{s_1}(t_2)+\alpha_{s_1}(t_3)),s_1\circ (s_2+s_3)-b(t_1\circ\alpha_{s_1}(t_2)-t_1,t_1)\\
        &&\qquad+b(t_1\circ\alpha_{s_1}(t_2)-t_1,t_1\circ\alpha_{s_1}(t_3)))\\
        &=&(t_1\circ(\alpha_{s_1}(t_2+t_3)),s_1\circ (s_2+s_3)+b(t_1\circ\alpha_{s_1}(t_2),t_1\circ\alpha_{s_1}(t_3))\\
        &&\qquad-b(t_1\circ\alpha_{s_1}(t_2)-t_1+t_1\circ\alpha_{s_1}(t_3),t_1))\\
        &=&(t_1\circ(\alpha_{s_1}(t_2+t_3)),s_1\circ (s_2+s_3)+b(t_1\circ\alpha_{s_1}(t_2),t_1\circ\alpha_{s_1}(t_3))\\
        &&\qquad-b(t_1\circ(\alpha_{s_1}(t_2)+\alpha_{s_1}(t_3)),t_1))\\
        &=&(t_1\circ(\alpha_{s_1}(t_2+t_3)),s_1\circ (s_2+s_3)+b(t_1\circ\alpha_{s_1}(t_2),t_1\circ\alpha_{s_1}(t_3))\\
        &&\qquad-b(t_1\circ\alpha_{s_1}(t_2+t_3)),t_1))\\
        &=&(t_1\circ(\alpha_{s_1}(t_2+t_3)),s_1\circ (s_2+s_3)-s_1+s_1\circ b(t_2,t_3))\\
        &=&(t_1\circ(\alpha_{s_1}(t_2+t_3)),s_1\circ (s_2+s_3+b(t_2,t_3))\\
        &=&(t_1,s_1)\circ (t_2+t_3,s_2+s_3+b(t_2,t_3))\\
        &=&(t_1,s_1)\circ ((t_2,s_2)+(t_3,s_3)).
	\end{eqnarray*}	
Therefore the result follows.
\end{proof}

Note that the lambda map of $T\rtimes_\circ S$ is defined by
\begin{eqnarray} \label{deflambda}
	\lambda_{(t_1,s_1)}(t_2,s_2)&=&
	\left(
	\lambda_{t_1}\alpha_{s_1}(t_2),~
	\lambda_{s_1}(s_2)-b(\lambda_{t_1}\alpha_{s_1}(t_2),t_1)
	\right),
\end{eqnarray}
and its socle is
$$
\Soc(T\rtimes_\circ S)=\{(t,s) : \lambda_s=\id_S,\;\lambda_t\circ
\alpha_s=\id_T,\; b(t,t_1)=0\mbox{ for all }t_1\in T\}.
$$
Moreover, the subset $T\times\{0\}$ is a normal subgroup of $(T\rtimes_\circ S,\circ)$, and $\{0\}\times S$ is a left
ideal of $T\rtimes_\circ S$.

A particular case of this theorem is when we assume that $b$ is a symmetric bilinear form. In this case, $b$ is a $2$-cocycle and
condition (\ref{cond3}) becomes
$$
\lambda_s(b(t_2,~t_3))=b(\lambda_{t_1}\alpha_{s}(t_2),~\lambda_{t_1}\alpha_{s}(t_3)),
$$
which is equivalent to the two conditions:
\begin{align}
	\lambda_s(b(t_2,~t_3))&=b(\alpha_{s}(t_2),~\alpha_{s}(t_3)),\label{condAlpha} \\
	b(t_2,~t_3)&=b(\lambda_{t_1}(t_2),~\lambda_{t_1}(t_3)),\label{condLambda}
\end{align}
for all $s\in S$ and $t_1,t_2,t_3\in T$.
Furthermore, if $T$ and $S$ are trivial skew braces of abelian type, then these two conditions (\ref{condAlpha}) and (\ref{condLambda}) reduce to 
$$b(t_2,~t_3)=b(\alpha_{s}(t_2),~\alpha_{s}(t_3)),$$
for all $s\in S$ and $t_2,t_3\in T$, i.e. $\alpha_s$ is in the orthogonal group of the bilinear form $b$, for all $s\in S$.
   
\section{Construction of simple braces of abelian type}

Let $p,q$ be two distinct primes. Consider the trivial skew braces over the abelian groups $T=(\Z/(p))^{n_1}\times (\Z/(q))^{n_2}$ and $S=\Z/(p)\times \Z/(q)$. Let $b_1\colon (\Z/(p))^{n_1}\times (\Z/(p))^{n_1}\rightarrow \Z/(p)$ and $b_2\colon (\Z/(q))^{n_2}\times (\Z/(q))^{n_2}\rightarrow \Z/(q)$ be two non-degenerate symmetric bilinear forms. Let $c_1$ be an element of order $q$ in the orthogonal group of $b_1$ and let $c_2$ be an element of order $p$ in the orthogonal group of $b_2$. Let $\alpha\colon S\rightarrow \Aut(T)$ be the map defined by $\alpha(i,j)=\alpha_{(i,j)}$ and $\alpha_{(i,j)}(u,v)=(c_1^{j}(u),c_2^{i}(v))$, for all $(i,j)\in S$ and $(u,v)\in T$. Let $b\colon T\times T\rightarrow S$ be the map defined by $b((u_1,v_1),(u_2,v_2))=(b_1(u_1,u_2),b_2(v_1,v_2))$, for all $(u_1,v_1),(u_2,v_2)\in T$. Then
\begin{eqnarray*}b(\alpha_{(i,j)}(u_1,v_1),\alpha_{(i,j)}(u_2,v_2))&=&(b_1(c_1^{j}(u_1),c_1^{j}(u_2)),b_2(c_2^{i}(v_1),c_2^{i}(v_2)))\\
	&=&(b_1(u_1,u_2),b_2(v_1,v_2)).
\end{eqnarray*}
Hence we can consider the asymmetric product $T\rtimes_{\circ}S$ of $T$ by $S$ (via $b$ and $\alpha$). 

\begin{theorem}\label{thm:simplebrace}
	With the above notation, the skew brace $T\rtimes_{\circ}S$ is simple if and only if $c_1-\id$ and $c_2-\id$ are invertible. 
\end{theorem}  

\begin{proof}
	Let $J=\{((u,v),(i,j)) : u\in\im(c_1-\id),\; v\in \im(c_2-\id), \; (i,j)\in S\}$. We shall prove that $J$ is an ideal of $T\rtimes_{\circ}S$. Clearly $J$ is a subgroup of $(T\rtimes_{\circ}S, +)$. Let $((u_1,v_1),(i_1,j_1))\in J$ and $((u_2,v_2),(i_2,j_2))\in T\rtimes_{\circ}S$. We have that
	\begin{eqnarray*}
		\lefteqn{\lambda_{((u_2,v_2),(i_2,j_2))}((u_1,v_1),(i_1,j_1))}\\
		&=&-((u_2,v_2),(i_2,j_2))+((u_2,v_2),(i_2,j_2))\circ ((u_1,v_1),(i_1,j_1))\\
		&=&((c_1^{j_2}(u_1),c_2^{i_2}(v_1)),(i_1-b_1(c_1^{j_2}(u_1),u_2),j_1-b_2(c_2^{i_2}(v_1),v_2)))\in J.
	\end{eqnarray*}
Hence $J$ is a left ideal of $T\rtimes_{\circ}S$. Furthermore
\begin{eqnarray*}
	\lefteqn{((u_1,v_1),(i_1,j_1))*((u_2,v_2),(i_2,j_2))}\\
	&=&-((u_1,v_1),(i_1,j_1))+((u_1,v_1),(i_1,j_1))\circ ((u_2,v_2),(i_2,j_2))-((u_2,v_2),(i_2,j_2))\\
	&=&((c_1^{j_1}(u_2),c_2^{i_1}(v_2)),(i_2-b_1(c_1^{j_1}(u_2),u_1),j_2-b_2(c_2^{i_1i_2}(v_2),v_1)))-((u_2,v_2),(i_2,j_2))\\
	&=&((c_1^{j_1}(u_2)-u_2,c_2^{i_1}(v_2)-v_2),(-b_1(c_1^{j_1}(u_2),u_1)-b_1(c_1^{j_1}(u_2)-u_2,u_2),-b_2(c_2^{i_1i_2}(v_2),v_2)\\
	&&\qquad\qquad -b_2(c_2^{i_1}(v_2)-v_2,v_2)))\in J.
\end{eqnarray*}	
Thus $J$ is an ideal of $T\rtimes_{\circ}S$. Therefore if $T\rtimes_{\circ}S$ is simple, then $c_1-\id$ and $c_2-\id$ are invertible.

Conversely, suppose that $c_1-\id$ and $c_2-\id$ are invertible. Let $I$ be a nonzero ideal of $T\rtimes_{\circ}S$.
Let $((u_1,v_1),(i_1,j_1))\in I$ be a nonzero element. Suppose first that $(u_1,v_1)=(0,0)$. In this case, $(i_1,j_1)\neq (0,0)$. We may assume that $i_1\neq 0$. Note that $q((0,0),(i_1,j_1))=((0,0),(qi_1,0))\in I$. Hence we may assume that $j_1=0$. Let $m$ be a positive integer such that $mi_1=1\in\Z/(p)$. Then
$m((0,0),(i_1,0))=((0,0),(1,0))\in I$. Now,  
\begin{eqnarray*}\lefteqn{((0,0),(1,0))*((u_2,v_2),(i_2,j_2))}\\
	&=&((0,(c_2-\id)(v_2)),(0,-b_2((c_2-\id)(v_2),v_2)))\in I,
\end{eqnarray*}
for all $((u_2,v_2),(i_2,j_2))\in T\rtimes_{\circ}S$. Since $c_2-\id$ is invertible, for every $w\in (\Z/(q))^{n_2}$ there exists $l_{w}\in \Z/(q)$ such that
$((0,w),(0,l_{w}))\in I$. Let $l\in \Z/(q)$. Note that if $w\neq 0$, since $b_2$ is non-degenerate, 
there exists $t\in (\Z/(q))^{n_2}$ such that $b_2(w,t)=l_{w}-l$. Hence $\lambda_{((0,t),(0,0))}((0,w),(0,l_w))=((0,w),(0,l_w-b_2(w,t)))=((0,w),(0,l))\in I$, for all $w\in (\Z/(q))^{n_2}\setminus\{0\}$ and $l\in \Z/(q)$. Thus, if $w\neq 0$, then $((0,w),(0,l))-((0,w),(0,0))=((0,0),(0,l))\in I$, for all $l\in\Z/(q)$. In particular, $((0,0),(0,1))\in I$. Now by a similar argument one can see that $((u,0),(k,0))\in I$, for all $u\in (\Z/(p))^{n_1}$ and $k\in\Z/(p)$. Now
$((u,0),(k,0))+((0,w),(0,l))=((u,w),(k,l))\in I$, for all $u\in (\Z/(p))^{n_1}$, $w\in (\Z/(q))^{n_2}$, $k\in\Z/(p)$ and $l\in\Z/(q)$.
Thus $I=T\rtimes_{\circ}S$ in this case. 

Suppose now that $(u_1,v_1)\neq (0,0)$. In this case we may assume that $u_1\neq 0$. Then $q((u_1,v_1),(i_1,j_1))=((qu_1,0),(i_2,j_2))\in I$, for some $(i_2,j_2)\in S$. Furthermore 
$$q((qu_1,0),(i_2,j_2))=((q^2u_1,0),(i_3,0))\in I,$$ 
for some $i_3\in\Z/(p)$. Since $q^2u_1\neq 0$ and $b_1$ is non-degenerate, there exists $u\in (\Z/(p))^{n_1}$ such that $b_1(q^2u_1,u)=i_3-k$. Hence
$\lambda_{((u,0),(0,0))}((q^2u_1,0),(i_3,0))=((q^2u_1,0),(i_3-b_1(q^2u_1,u)))=((q^2u_1,0),(k,0))\in I$, for all $k\in\Z/(p)$, and then
$((q^2u_1,0),(1,0))-((q^1u_1,0),(0,0))=((0,0),(1,0))\in I$. Now, as above, one can prove that $I=T\rtimes_{\circ}S$, and the result follows.
\end{proof}

We shall give now some concrete constructions of simple skew braces of abelian type using Theorem \ref{thm:simplebrace}. Let $p,q$ be two distinct primes. Let $n,m$ be two positive integers. Let
$$D_1=\left(\begin{array}{rrrrr}
0&0&\ldots &0&-1\\
\vspace{-5pt}1&0&\ldots &0&-1\\
\vspace{-5pt}0&\ddots &\ddots &\vdots&\vdots\\
\vdots &\ddots &\ddots &0&-1\\
0&\ldots&0&1&-1\end{array}\right),\; E_1=\left(\begin{array}{cccc}
\vspace{-5pt}1-q&1&\ldots &1\\
\vspace{-5pt}1&\ddots &\ddots&\vdots\\
\vdots &\ddots &\ddots &1\\
1&\ldots&1&1-q\end{array}\right)\in M_{q-1}(\Z/(p))$$
 and  
 $$D_2=\left(\begin{array}{rrrrr}
0&0&\ldots&0&-1\\
\vspace{-5pt}1&0&\ldots&0&-1\\
\vspace{-5pt}0&\ddots&\ddots&\vdots&\vdots\\
\vdots&\ddots&\ddots&0&-1\\
0&\ldots&0&1&-1\end{array}\right),\; E_2=\left(\begin{array}{cccc}
\vspace{-5pt}1-p&1&\ldots &1\\
\vspace{-5pt}1&\ddots &\ddots&\vdots\\
\vdots &\ddots &\ddots &1\\
1&\ldots&1&1-p\end{array}\right)\in M_{p-1}(\Z/(q)).$$ 
Note that $D_1$ is the companion matrix of the polynomial $1+x+\dots +x^{q-1}$ and thus it has order $q$. Similarly $D_2$ has order $p$. 
It is easy to check that $D_1^tE_1D_1=E_1$ and $D_2^tE_2D_2=E_2$. Consider in $M_{n_1(q-1)}(\Z/(p))$ the block diagonal matrices with $n_1$ blocks of degree $q-1$:
$$C_1=\left(\begin{array}{cccc}
	D_1&0&\ldots &0\\
	0&\ddots &\ddots&\vdots\\
	\vdots &\ddots &\ddots &0\\
	0&\ldots&0&D_1\end{array}\right)\mbox{ and } B_1=\left(\begin{array}{cccc}
	E_1&0&\ldots &0\\
	0&\ddots &\ddots&\vdots\\
	\vdots &\ddots &\ddots &0\\
	0&\ldots&0&E_1\end{array}\right).$$
Similarly consider in $M_{n_2(p-1)}(\Z/(q))$ the block diagonal matrices with $n_2$ blocks of degree $p-1$:
$$C_2=\left(\begin{array}{cccc}
	D_2&0&\ldots &0\\
	0&\ddots &\ddots&\vdots\\
	\vdots &\ddots &\ddots &0\\
	0&\ldots&0&D_2\end{array}\right)\mbox{ and } B_2=\left(\begin{array}{cccc}
	E_2&0&\ldots &0\\
	0&\ddots &\ddots&\vdots\\
	\vdots &\ddots &\ddots &0\\
	0&\ldots&0&E_2\end{array}\right).$$
Let $T=(\Z/(p))^{n_1(q-1)}\times (\Z/(q))^{n_2(p-1)}$ and $S=\Z/(p)\times\Z/(q)$ be the trivial skew braces over these abelian groups.
Let $b\colon T\times T\rightarrow S$ be the symmetric bilinear map defined by
$$b((u_1,v_1),(u_2,v_2))=(u_1B_1u_2^t,v_1B_2v_2^t),$$
for all $u_1,u_2\in (\Z/(p))^{n_1(q-1)}$ and $v_1,v_2\in (\Z/(q))^{n_2(p-1)}$. Note that $\det(B_1)=(-q)^{q-2}\neq 0$ and $\det(B_2)=(-p)^{p-2}\neq 0$, thus the symmetric bilinear forms associated to these matrices are non-degenerate.
Let $\alpha\colon S\rightarrow \Aut(T)$ be the map defined by $\alpha(i,j)=\alpha_{(i,j)}$ and $\alpha_{(i,j)}(u,v)=(u(C_1^{t})^j,v(C_2^t)^{i})$, for all $(i,j)\in S$ and $(u,v)\in T$. Since $C_k^tB_kC_k=B_k$, for $k=1,2$, we can consider the asymmetric product $T\rtimes_{\circ}S$ of $T$ by $S$ (via $b$ and $\alpha$). Since $C_k-I$, where $I$ is the identity matrix, is invertible, by Theorem \ref{thm:simplebrace}, $T\rtimes_{\circ}S$ is a simple skew brace of abelian type. 

Another construction is based in the following fact. Let $D$ be an invertible matrix in $M_n(K)$, where $K$ is a field. Consider the block matrices
$$C=\left(\begin{array}{cc}
D^t&0\\
0&D^{-1}\end{array}\right)\in M_{2n}(K)\quad\mbox{and}\quad B=\left(\begin{array}{cc}
0&I_n\\
I_n&0\end{array}\right)\in M_{2n}(K).$$
Then one can check that $C^tBC=B$.     
Thus we can take any invertible matrix $D_1\in M_{m_1}(\Z/(p))$ of order $q$ and any matrix $D_2\in M_{m_2}(\Z/(q))$ of order $p$, such that $D_k-I$ also is invertible for $k=1,2$. 
Consider in $M_{2n_1m_1}(\Z/(p))$ the block square matrices with $2n_1$ blocks of degree $m_1$ in the diagonal:
$$C_1=\left(\begin{array}{cccccc}
	D_1^t&0&0&\ldots &0&0\\
	0&D_1^{-1}&0&\ldots &0&0\\
	0&0&\ddots &\ddots &\vdots&\vdots \\
	\vdots&\vdots&\ddots&\ddots&0&0\\
	0&0&\ldots&0&D_1^t&0\\
    0&0&\ldots&0&0&D_1^{-1}\end{array}\right)\mbox{ and } B_1=\left(\begin{array}{cccccc}
	0&I_{m_1}&0&\ldots &0&0\\
	I_{m_1}&0&0&\ldots &0&0\\
	0&0&\ddots&\ddots&\vdots&\vdots\\
	\vdots&\vdots &\ddots &\ddots &0&0\\
	0&0&\ldots&0&0&I_{m_1}\\
	0&0&\ldots&0&I_{m_1}&0
\end{array}\right).$$
Similarly consider in $M_{2n_2m_2}(\Z/(q))$ the block square matrices with $2n_2$ blocks of degree $m_2$ in the diagonal:
$$C_2=\left(\begin{array}{cccccc}
	D_2^t&0&0&\ldots &0&0\\
	0&D_2^{-1}&0&\ldots &0&0\\
	0&0&\ddots &\ddots &\vdots&\vdots \\
	\vdots&\vdots&\ddots&\ddots&0&0\\
	0&0&\ldots&0&D_2^t&0\\
	0&0&\ldots&0&0&D_2^{-1}\end{array}\right)\mbox{ and } B_1=\left(\begin{array}{cccccc}
	0&I_{m_2}&0&\ldots &0&0\\
	I_{m_2}&0&0&\ldots &0&0\\
	0&0&\ddots&\ddots&\vdots&\vdots\\
	\vdots&\vdots &\ddots &\ddots &0&0\\
	0&0&\ldots&0&0&I_{m_2}\\
	0&0&\ldots&0&I_{m_2}&0
\end{array}\right).$$
Let $T=(\Z/(p))^{2n_1m_1}\times (\Z/(q))^{2n_2m_2}$ and $S=\Z/(p)\times\Z/(q)$ be the trivial skew braces over these abelian groups.
Let $b\colon T\times T\rightarrow S$ be the symmetric bilinear map defined by
$$b((u_1,v_1),(u_2,v_2))=(u_1B_1u_2^t,v_1B_2v_2^t),$$
for all $u_1,u_2\in (\Z/(p))^{2n_1m_1}$ and $v_1,v_2\in (\Z/(q))^{2n_2m_2}$.
Let $\alpha\colon S\rightarrow \Aut(T)$ be the map defined by $\alpha(i,j)=\alpha_{(i,j)}$ and $\alpha_{(i,j)}(u,v)=(u(C_1^{t})^j,v(C_2^t)^{i})$, for all $(i,j)\in S$ and $(u,v)\in T$. Since $C_k^tB_kC_k=B_k$, for $k=1,2$, we can consider the asymmetric product $T\rtimes_{\circ}S$ of $T$ by $S$ (via $b$ and $\alpha$). Since $C_k-I$, where $I$ is the identity matrix, is invertible, by Theorem \ref{thm:simplebrace}, $T\rtimes_{\circ}S$ is a simple skew brace of abelian type. 

For example, for $p=2$ and $q=31$, with the first construction we obtain simple skew braces of abelian type of order $2^{30n_1}31^{n_2}$ for all positive integers $n_1,n_2$. With the second construction, taking
$$D_1=\left(\begin{array}{ccccc}
0&0&0&0&1\\
1&0&0&0&0\\
0&1&0&0&1\\
0&0&1&0&0\\
0&0&0&1&0\end{array}\right)\in M_5(\Z/(2)),$$
we obtain simple skew braces of abelian type of order $2^{10n_1}31^{2n_2}$ for all positive integers $n_1,n_2$. Using the two methods one can easily construct simple skew braces of abelian type of order $2^{10n_1}31^{n_2}$ for all positive integers $n_1,n_2$.



\section{Exercises}

\begin{prob} Construct a concret example of a simple skew brace of abelian type of order $2^{10}31$.
\end{prob}

\begin{prob}
\end{prob}


\begin{prob}
\end{prob}

\begin{prob}
\end{prob}

\section{Open problems}

\begin{problem}
Is every finite simple skew brace of abelian type the asymmetric product of two nonzero skew braces?
\end{problem}

\section{Notes}

\index{Bachiller, D.}
\index{Ced\'o, F}
\index{Jespers, E.}
\index{Okni\'{n}ski, J.}
The first example of a simple non-trivial  skew brace of abelian type is due to Bachiller \cite{MR3763276}. 
In the same papar Bachiller introduced the matched product of two skew braces of abelian type. One can see that the definition of matched product of two skew braces is the same that in the case of skew braces of abelian type. 
In \cite{MR3812099}, Bachiller, Ced\'o, Jespers and Okni\'{n}ski introduced the matched product of more than two skew braces of abelian type (this can be generaliced to arbitrari skew braces without changes). In the same paper the authors construct several families of  simple non-trivial skew braces of abelian type using matched products.

\index{Catino, F.}
\index{Colazzo, I.}
\index{Stefanelli, P.}
The asymmetric product of skew braces of abelian type was intruduced by Catino, Colazzo and Stefanelli in \cite{MR3478858}, and Theorem \ref{ccs} is due to them (\cite[Theorem 3]{MR3478858}).  In \cite{MR4020748} Bachiller, Ced\'o, Jespers and Okni\'{n}ski used the asymmetric product to construct  new families of simple non-trivial skew braces of abelian type. In fact, every known simple non-trivial skew brace is an asymmetric product (see \cite{MR4020748, MR4161288, MR4122077}).
Theorem \ref{thm:simplebrace} is based on \cite[Theorem 3.6]{MR3812099} and \cite[Theorem 6.2]{MR4020748}. The two concrete constructions of simple non-trivial simple skew braces of abelian type appear in \cite{MR3812099,  MR4161288}.
